---
title: "Functions for querying the HGNC REST API"
author:
- name: Marcel Ramos
  affiliation: CUNY School of Public Health, New York, NY
date: "`r BiocStyle::doc_date()`"
vignette: |
  %\VignetteIndexEntry{HGNCREST}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    toc_float: true
package: HGNCREST
---

# HGNCREST

The `HGNCREST` package provides functions for querying the HGNC REST API. The
functions follow the HUGO Gene Nomenclature Committee (HGNC) REST API
documentation at <https://www.genenames.org/help/rest/>. There are three main
operations that can be performed with this package:

1. fetching general information about the HGNC database (`hgnc_info`).
2. fetching information about a specific gene (`hgnc_fetch`)
3. searching for genes based on a query (`hgnc_search`)

# Installation

```{r, eval = FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("waldronlab/HGNCREST")
```

## Package load

```{r,include=TRUE,results="hide",message=FALSE,warning=FALSE}
library(HGNCREST)
```

# General information

The `hgnc_info` function returns general information about the HGNC database.
It includes `searchableFields` and `storedFields` metadata.

```{r}
hgnc_info()
```

# Searchable fields

The `searchableFields` function is a convenience function that returns a
character vector of searchable fields in the HGNC database.

```{r}
searchableFields()
```

# Fetching gene information

The `hgnc_fetch` function returns a `tibble` with information about the gene
specified by the `searchableField` and `value` arguments.

```{r}
hgnc_fetch("ena", "BC040926")
```

# Searching for genes

The `hgnc_search` function searches for genes based on a query. It returns a
`data.frame` with information about the genes that match the query.

```{r}
hgnc_search("symbol", "BRCA1")
```

# Conclusion

The `HGNCREST` package provides a convenient way to query the HGNC REST API
from R. It allows users to fetch information about specific genes, search for
genes based on a query, and get general information about the HGNC database.

# sessionInfo

<details> <summary> Click to expand </summary>

```{r}
sessionInfo()
```

</details>
